---
title: "Reproducible Research: Peer Assessment 1"

output: md_document
        
html_document: PA1_template.html
        
keep_md: true

---
        
## Loading and preprocessing the data

At the very begining, we need to check the existence of files in a certain directory. Here come the codes.
```{r checkfiles, cache=TRUE}
wd <- getwd()
 if(!file.exists("./data")){
+     dir.create("./data")  
}
if(!file.exists("./data/repdata_data_activity.zip")){
+     url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
+     download.file(url, destfile="./data/repdata_data_activity.zip", method="curl")
}
 if(!file.exists("./data/activity.csv")){
+     zipF<-file.choose() 
+     outDir<-"./data" 
+     unzip(zipF,exdir=outDir)
} 
```

Now, let's load the data! 
```{r loaddata}
setwd("./data")
rawdata <- read.csv("activity.csv")
```

To preprocess the data set, here we only need to change the format of date *from factor to date*.
```{r preprocess}
#Take a glance first
summary(rawdata)
str(rawdata)
#preprocess it
mydata <- rawdata
mydata$date <- as.Date(mydata$date)
```
        
## What is mean total number of steps taken per day?

Before following calculations, we need to drop all the missing values in the data set. 
```{r dropNA}
nona <- mydata[complete.cases(mydata),]
str(nona)
```


Now, calculate the total number of steps taken per day. 
```{r sumsteps}
#Create a new dataframe to store the total number of steps each day
totalsteps <- aggregate(nona$steps, by=list(Date=nona$date), FUN=sum)
names(totalsteps)<-c("Date","Steps")
totalsteps
```


Then, make a histgram of the total number of steps taken each day.
```{r histgram}
hist(totalsteps$Steps, main="Total Number of Steps taken Each Day", xlab="Steps", ylim= c(0,30))
```


Calculate and report the mean and median of the total number of steps taken each day.
```{r mean&med, results='hide'}
#Calculate mean and median
mean <- mean(totalsteps$Steps)
median <- median(totalsteps$Steps)
```

So the mean of the total number of steps is `r mean` and the median of that is `r median`

## What is the average daily activity pattern?

To understand the average daily activity pattern, we coudl first make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
```{r timeplot}
#calculate the average number of steps taken
pattern <- aggregate(nona$steps, by=list(Intervals = nona$interval),FUN=mean)
names(pattern) <- c("Intervals","AvgSteps")
#make a plot
with(pattern, plot(Intervals, AvgSteps, type="n", ylab="Average Number of Steps taken", xlab="Intervals",main="Time Series Plot"))
with(pattern, lines(Intervals, AvgSteps, type="l",col="red"))
```

Also, we could tell which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps.
```{r maxnum}
maxium <- pattern$Intervals[which(pattern$AvgSteps==max(pattern$AvgSteps))]
```

The 5-minute interval which contains the maximum number of steps is `r maxium`.


## Imputing missing values

Before imputing missing values, we could calculate and report the  total number of missing values in the dataset (i.e. the total number of rows with NA)

```{r countNA}
totalNA <- sum(is.na(mydata$steps))
```

The  total number of missing values in the dataset is `r totalNA`.

Then, we could create a newdataset **newdata** with all of the missing values filled in. Here we use **the mean for that 5-minute interval**.
```{r fillNA}
newdata <- mydata
na <- is.na(newdata$steps)
newdata[na, ]$steps <- mean(pattern$AvgSteps)
```

After that, we could make the histogram of the total number of steps taken each day and the mean and median total number of steps taken per day; then, compare these values with the first part of this assignment to see the difference. 
```{r compare}
#make a new histgram
totalsteps2 <- aggregate(newdata$steps, by=list(Date=newdata$date), FUN=sum)
names(totalsteps2)<-c("Date","Steps")
par(mfrow=c(1,2),mar=c(4,4,2,1),cex.main=0.6)
hist(totalsteps$Steps, main="Total Number of Steps taken Each Day", xlab="Steps", ylim= c(0,40))
hist(totalsteps2$Steps, main="Total Number of Steps taken Each Day(Revised)", xlab="Steps", ylim= c(0,40))
#recalculate the mean and the median
mean2 <- mean(totalsteps2$Steps)
median2 <- median(totalsteps2$Steps)
```

The **new** mean of the total number of steps is `r mean2` and the **new** median of that is `r median2`, while the **former** mean  is `r mean` and the **former** median  is `r median`.


By filling in missing values with mean for that 5-minute interval, we could see clearly that **the frequency of total number of steps around 10,000 increases while other parts remain the same**, and that **the mean remains the same while the meidan change from `r median` to the `r median2`, exactly the mean**. 


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - ¡°weekday¡± and ¡°weekend¡± indicating whether a given date is a weekday or weekend day.
```{r weekday, results='hide'}
Sys.setlocale("LC_TIME","US")
weekdata <- newdata
weekdata$weekday <- weekdays(weekdata$date)
weekdata$type <- with(weekdata, ifelse(weekday=="Saturday"|weekday=="Sunday","weekend","weekday"))
weekdata$type <- as.factor(weekdata$type)
```

Finally, make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r plot2}
pattern2 <- aggregate(weekdata$steps, by=list(Intervals = weekdata$interval, Type=weekdata$type),FUN=mean)
names(pattern2) <- c("Intervals","Type","AvgSteps")
library(lattice)
xyplot(AvgSteps~Intervals|Type,data=pattern2,layout=c(1,2),type="l",ylab="Number of Steps")
```

END